---
title: "p8105_hw1_mz2777"
author: "Mengyu Zhang"
date: "9/14/2019"
output: github_document
---

# Problem 1

Create the data frame.

```{r, echo = FALSE, include = FALSE}
library(tidyverse)
```

```{r create the data frame}
#Create the data frame
df_p1 = tibble(
  norm_samp = rnorm(8, mean = 0, sd = 1),
  norm_samp_logical = norm_samp > 0,
  char_vec = c("A", "B", "C", "A", "B", "B", "C", "B"),
  fac_vec = factor(char_vec, labels = c("I","II","III"))
)
```


Character and factor variables did not work when the `mean` function was applied to them. The messages are as follows.

```{r take the mean, echo = FALSE, message = TRUE}
#Take the mean of each variable in the data frame
mean_samp = mean(pull(df_p1, norm_samp))
mean_samp_logical = mean(pull(df_p1, norm_samp_logical))
mean_char_vec = mean(pull(df_p1, char_vec))
mean_fac_vec = mean(pull(df_p1, fac_vec))
```

After applying the `as.numeric` function to the logical, character and factor variables, logical and factor variables are able to be converted to numeric variable. Since unlike character variables, these two kind of variables can be represented by the numbers without losing their logic and meanings. Therefore, this conversion can help explain why the mean value of the character variables can not be calculated.

```{r, eval = FALSE}
as.numeric(pull(df_p1, norm_samp_logical))
as.numeric(pull(df_p1, char_vec))
as.numeric(pull(df_p1, fac_vec))
```

Here are some conversions of variables.

```{r }
as.numeric(pull(df_p1, norm_samp_logical)) * pull(df_p1, norm_samp)
factor(pull(df_p1, norm_samp_logical)) * pull(df_p1, norm_samp)
as.numeric(factor(pull(df_p1, norm_samp_logical))) * pull(df_p1, norm_samp)
```

# Problem 2

```{r}
df_p2 = tibble(
  x = rnorm(500, mean = 0, sd = 1),
  y = rnorm(500, mean = 0, sd = 1),
  logic_vec = x + y > 1,
  logic2num = as.numeric(logic_vec),
  logic2fac = factor(logic_vec)
)
```

A short description of the data frame I have built.

* The size of the dataset is `r nrow(df_p2)`*`r ncol(df_p2)`.
* The mean of x is `r mean(df_p2$x)`.
* The median of x is `r median(df_p2$x)`.
* The standard deviation of x is `r sd(df_p2$x)`.
* The proportion of cases for which x + y > 1 is `r mean(pull(df_p2, logic2num))*100`%.

Make a scatterplot of y vs x; color points using the logical variable. Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and **comment on the color scales**.

```{r}
pic1 = ggplot(df_p2, aes(x = x, y = y, color = logic_vec)) + geom_point()
pic2 = ggplot(df_p2, aes(x = x, y = y, color = logic2num)) + geom_point()
pic3 = ggplot(df_p2, aes(x = x, y = y, color = logic2fac)) + geom_point()
pic1
pic2
pic3
ggsave("scatter point.png", plot = pic1)
```


