---
title: "p8105_hw1_mz2777"
author: "Mengyu Zhang"
date: "9/14/2019"
output: github_document
---

# Problem 1

### Create the Dataframe

```{r, echo = FALSE, include = FALSE}
library(tidyverse)
```

```{r create the data frame}
#Create the data frame
df_p1 = tibble(
  norm_samp = rnorm(8, mean = 0, sd = 1),
  norm_samp_logical = norm_samp > 0,
  char_vec = c("A", "B", "C", "A", "B", "B", "C", "B"),
  fac_vec = factor(char_vec, labels = c("I","II","III"))
)
```

### Take the Mean
Character and factor variables did not work when the `mean` function was applied to them. The erroe messages are as follows.

```{r take the mean, echo = FALSE, message = TRUE}
#Take the mean of each variable in the data frame
mean_samp = mean(pull(df_p1, norm_samp))
mean_samp_logical = mean(pull(df_p1, norm_samp_logical))
mean_char_vec = mean(pull(df_p1, char_vec))
mean_fac_vec = mean(pull(df_p1, fac_vec))
```

### Applying the `as.numeric` Function
After applying the `as.numeric` function to the logical, character and factor variables, logical and factor variables are able to be converted to numeric variable. Since unlike character variables, these two kind of variables can be represented by the numbers without losing their logic and meanings. Therefore, this conversion can help explain why the mean value of the character variables can not be calculated.

```{r, eval = FALSE}
#as.numeric
as.numeric(pull(df_p1, norm_samp_logical))
as.numeric(pull(df_p1, char_vec))
as.numeric(pull(df_p1, fac_vec))
```

### Conversions and Multiplications.
* convert the logical vector to numeric, and multiply the random sample by the result
* convert the logical vector to a factor, and multiply the random sample by the result
* convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result

```{r }
#convert the logical vector to numeric, and multiply the random sample by the result
as.numeric(pull(df_p1, norm_samp_logical)) * pull(df_p1, norm_samp)
#convert the logical vector to a factor, and multiply the random sample by the result
factor(pull(df_p1, norm_samp_logical)) * pull(df_p1, norm_samp)
#convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result
as.numeric(factor(pull(df_p1, norm_samp_logical))) * pull(df_p1, norm_samp)
```

# Problem 2

### Create the Dataframe
```{r}
#create the dataframe
df_p2 = tibble(
  x = rnorm(500, mean = 0, sd = 1),
  y = rnorm(500, mean = 0, sd = 1),
  logic_vec = x + y > 1,
  logic2num = as.numeric(logic_vec),
  logic2fac = factor(logic_vec)
)
```

### Short Description

A short description of the data frame I have built.

* The size of the dataset is `r nrow(df_p2)`*`r ncol(df_p2)`.
* The mean of x is `r mean(df_p2$x)`.
* The median of x is `r median(df_p2$x)`.
* The standard deviation of x is `r sd(df_p2$x)`.
* The proportion of cases for which x + y > 1 is `r mean(pull(df_p2, logic2num))*100`%.

### Scatterplot

Make scatterplots of y vs x that color points using the logical, numeric and factor variables, respectively.

#### Color Scale
We can see from legends that the logical and factor variables are binary, true or false, and there are only two kind of colors in the legends. However, for numeric variable, the legend's rage is from 0 to 1 though there is no value between 0 and 1.

```{r}
#color point using the logical variable
pic1 = ggplot(df_p2, aes(x = x, y = y, color = logic_vec)) + 
  geom_point() + ggtitle("Color points using the logical variable")
pic1

#color point using the numeric variable
pic2 = ggplot(df_p2, aes(x = x, y = y, color = logic2num)) + 
  geom_point() + ggtitle("Color points using the numeric variable")
pic2

#color point using the factor variable
pic3 = ggplot(df_p2, aes(x = x, y = y, color = logic2fac)) + 
  geom_point() + ggtitle("Color points using the factor variable")
pic3

#export the first scatterplot
ggsave("scatter point.png", plot = pic1)
```

